---
title: "Statistics for the manuscript"
format: html
editor: visual
---

## Libraries

```{r}
#| warning: false
#| message: false
library(data.table)
library(tidyverse)
library(huxtable)
source("../analysis/code/common_utils.R")
output_for_paper_dir = "."
```

## Results, SBS subsection

```{r}
fread(
  file.path(output_for_paper_dir,
            "summary_stats_SBS.csv")) %>%
  filter(Tool %in% global_raw_tools_to_plot) %>%
  arrange(desc(m.Combined)) %>% 
  head() %>% select(Tool, m.Combined)

sbs_file <- file.path(output_for_paper_dir, 
                      "assessment_each_sample_SBS.csv")

sbs_assessment <- data.table::fread(sbs_file)
```

#### Results, SBS subsection, page 18

"The Combined Scores of PASA and MuSiCal across all 900 synthetic SBS spectra were similar (means 2.64 and 2.62, respectively, p = 0.072, 2-sided Wilcoxon rank-sum test). "

```{r}
wilcox.test(x = sbs_assessment[Tool == "pasa",]$Combined, 
              y = sbs_assessment[Tool == "musical",]$Combined) %>% `$`(p.value)
```

```{r}
wilcox.test(x = sbs_assessment[Tool == "pasa",]$Combined, 
            y = sbs_assessment[Tool == "fitms_0.010",]$Combined) %>% `$`(p.value)

wilcox.test(x = sbs_assessment[Tool == "musical",]$Combined, 
            y = sbs_assessment[Tool == "fitms_0.010",]$Combined) %>% `$`(p.value)

```

### MutationalPatterns SBS ranks

Pge 18, "As another example, the ranks of MutationalPatterns ranged from 4 (Kidney-RCC and Ovary-AdenoCA) to 12 (Skin-Melanoma). "

```{r}
sbs_tool_rank = 
fread("tools_ranked_in_each_cancer_type_SBS.csv")

# To inspect the data:
filter(sbs_tool_rank, Tool == "MutPat") %>%
  select(`Cancer type`, Tool,`Rank within cancer type`, `mean (Combined Score)`)

sbs_tool_rank %>%
  group_by(Tool) %>%
 summarize(
    min_rank = min(`Rank within cancer type`),
    max_rank = max(`Rank within cancer type`),
    difference = max_rank - min_rank
  ) 


```

### SBS: Worst cancer type for each tool

```{r}
#| eval: false
# Not used in paper; the .csv input has been replaced by an Excel file.

fread(
  file.path(
    output_for_paper_dir,
    "summary_stats_by_cancer_type_SBS.csv")) %>%
  select(Tool, cancer.type, m.Combined) %>%
  filter(Tool %in% global_raw_tools_to_plot) %>%
  group_by(Tool) %>% 
  mutate(rank_in_tool = min_rank(m.Combined)) %>%
  arrange(m.Combined, .by_group = TRUE) %>%
  dplyr::slice(1:2) %>% 
  filter(cancer.type == "Skin-Melanoma") %>%
  data.frame()

```

## Results, DBS subsection

Page 18, "On synthetic DBS spectra, PASA had the highest Combined Score, which was significantly higher than that of FitMS, which had the next highest (mean 2.78 versus 2.74, p\<9.9×10\^(-9), 2-sided Wilcoxon rank-sum test, Figure 4, S4, Table S12). The Combined Scores of the other approaches were much lower, with 3rd-ranked MuSiCal having a mean Combined Score of 2.57, significantly lower that of FitMS (p \< 3.9×10\^(-44), 2-sided Wilcoxon rank-sum test). "

```{r}
fread(
  file.path(output_for_paper_dir,
            "summary_stats_DBS.csv")) %>%
  arrange(desc(m.Combined)) %>%
  filter(Tool %in% global_raw_tools_to_plot) %>%
  head() %>% select(Tool, m.Combined)

dbs_file <- file.path(output_for_paper_dir, 
                      "assessment_each_sample_DBS.csv")
dbs_assessment <- fread(dbs_file)

wilcox.test(
  x = dbs_assessment[Tool == "pasa",]$Combined, 
  y = dbs_assessment[Tool == "fitms_0.010",]$Combined) %>%
  `$`(p.value)


wilcox.test(
  x = dbs_assessment[Tool == "fitms_0.010",]$Combined, 
  y = dbs_assessment[Tool == "musical",]$Combined) %>%
  `$`(p.value)


```

```{r}
dbs_tool_rank = 
fread("tools_ranked_in_each_cancer_type_DBS.csv")
dbs_tool_rank %>%
  group_by(Tool) %>%
 summarize(
    min_rank = min(`Rank within cancer type`),
    max_rank = max(`Rank within cancer type`),
    difference = max_rank - min_rank
  )

```

Page 20, "As another example, deconstructSigs and SigPro tied for 1st in Skin-Melanoma but were both ranked 12th in several cancer types."

```{r}
dbs_tool_rank %>% 
  filter(Tool %in% c("DeconSig", "SigPro"), `Rank within cancer type` %in% c(1,12))
```

### DBS: comparison for recall for each tool versus SBS

Page 20, paragraph starting "Recall (sensitivity) for DBS attributions was significantly better than for SBSs for 10 of the approaches ..."

```{r}
getwd()
sbss = select(sbs_assessment, Tool, Sample.ID, cancer.type, sens)
dbss = select(dbs_assessment, Tool, Sample.ID, cancer.type, sens)
inner_join(sbss, dbss, 
           by = join_by(Tool == Tool, 
                        cancer.type == cancer.type, 
                        Sample.ID == Sample.ID)) %>%
  mutate(SBS.recall = sens.x, DBS.recall = sens.y,
         .keep = "unused") %>%
  filter(Tool %in% tools_to_plot_and_order("DBS")) %>%
  group_by(Tool) %>%
  summarize(
    median_dif = median(DBS.recall - SBS.recall),
    mean_dif = mean(DBS.recall - SBS.recall),
    wilcox.test = list(
      broom::tidy(
        wilcox.test(jitter(SBS.recall),
                    DBS.recall, paired = TRUE)
      )
    )
  ) %>%
  unnest(cols = c(wilcox.test)) %>%
  mutate(method = NULL, 
         statistic = NULL, 
         alternative = NULL,
         FDR = p.adjust(p.value)) %>%
  filter(FDR < 0.1) ->foo
foo
```

```{r}
cat("Number of tools with better recall in DBS: ", nrow(filter(foo, mean_dif < 0)), "\n")

cat("Number of tools with worse recall in DBS: ", nrow(filter(foo, mean_dif > 0)))
 
```

## Results, ID subsection

Page 21, 1st paragrap of "Accuracy on ID...."

```{r}
fread(
  file.path(output_for_paper_dir,
            "summary_stats_ID.csv")) %>%
  filter(Tool %in% global_raw_tools_to_plot) %>%
  arrange(desc(m.Combined)) %>%
  head() %>% 
  select(Tool, m.Combined) %>%
  filter(Tool %in% c("pasa", "fitms_0.010", "musical"))

id_file <- file.path(output_for_paper_dir, 
                     "assessment_each_sample_ID.csv")
id_dt <- data.table::fread(id_file)

wilcox.test(x = id_dt[Tool == "pasa",]$Combined, 
            y = id_dt[Tool == "fitms_0.010",]$Combined) %>% `$`(p.value)

wilcox.test(x = id_dt[Tool == "fitms_0.010",]$Combined, 
            y = id_dt[Tool == "musical",]$Combined) %>% `$`(p.value)
```

```{r}
id_tool_rank = 
fread("tools_ranked_in_each_cancer_type_ID.csv")

id_tool_rank %>%
  filter(Tool == "PASA")
```

```{r}

id_tool_rank %>%
  group_by(Tool) %>%
 summarize(
    min_rank = min(`Rank within cancer type`),
    max_rank = max(`Rank within cancer type`),
    difference = max_rank - min_rank
  ) %>% as_huxtable() %>% huxtable::set_caption("ID: difference between min and max ranks")


```

```{r}
id_tool_rank %>% filter(Tool == "SigPro")
```
